### 2016-03-27 tests/HJM/manuscript.tex
working with a reasonable text

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{mathrsfs }
\usepackage[section]{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{patchplots}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{multirow}

%line suggested by my compiler (Fabian)
\pgfplotsset{compat=1.9}

% put figures at the end of the document, while I am writing is very annoying to have the figures in the middle of the text! 
% \usepackage[nomarkers,figuresonly]{endfloat}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\input{macros.tex}

\title[MIMC and MLMC HJM]{Multi-Index and Multi-Level Monte Carlo Evaluation of HJM Models}
\author{Juho H\"app\"ol\"a}
\address{\kaustaddress}
\email{juho.happola@iki.fi}

\date{\today}    
                                    % Activate to display a given date or no date
\begin{document}

\begin{abstract}
Abstract here!!!
\end{abstract}

\maketitle
%\tableofcontents

We are interested in analysing the Forward Euler Finite Element evaluation of forward curve models of HJM models. The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.

We are focused on the evolution of the forward curve $f \parent{t , \tau}$ for $t \in [0,t_{max}]$, $\tau \in [0,\tau_{max}]$ with $\tau_{max}>t_{max}$. In order to define a mesh for the numerical solution of the problem, let us define $\bar \ell \in \mathbb{N}^3$.

Let us first discretise the time interval as
\begin{align*}
\Delta t_{\bar \ell} &= 2^{-\ell_1} t_{max},
\\
t_{\bar \ell, n} &= n \Delta t_{\bar \ell}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_1}].
\end{align*}
Since there is an additional requirement that $t_{max}$ be in
the mesh of maturities, one point in the mesh in $\tau \in [0,\tau_{max}]$. We let $\ell_1$ and $\ell_2$ define the number of discretisation points to the right and to the left of the fixed
point, respectively.
\begin{align*}
\Delta \tau_{\bar \ell,1} &= 2^{-\ell_3} t_{max},  \\
\Delta \tau_{\bar \ell,2} &= 2^{-\ell_2} \parent{\tau_{max}-t_{max}}, \\
\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_3}] \\
\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [2^{\ell_3},2^{\ell_3}+1,\cdots, 2^{\ell_3}+2^{\ell 2}].
\end{align*}

Using this notation, we have the following rates of convergence, following the Multi-Index notation:
\begin{align}
\bar \beta &= [2,2,2]^T \\
\bar \gamma &= [1,1,1]^T .
\end{align} 
Choosing the Simpson method for integration, we have the order of quadrature $p_Q=3$, which brings the quadrature error to 
\begin{align}
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}} \leq C \parent{\Delta \tau_{\bar \ell}}^{2 \kappa p_Q} = 
\tilde C 2^{-6 \kappa \mathrm{min} \parent{\ell_3,\ell_2}}.
\end{align}
On the other hand, the error of the simpson method is the sum of errors for $\tau < t_{max}$ and $\tau \leq t_{max}$. The error for these two domains are $\bigo{2^{-2 \ell_2}}$ and $\bigo {2^{-2 \ell_3}}$, respectively. Combining together, we get the weak and the strong rate
for the Simpson quadrature:
\begin{align}
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2}} \leq&  \tilde C_Q 2^{-6 \mathrm{min} \parent{\ell_3,\ell_2}}
\\
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}} \leq& C_Q 2^{-2 \mathrm{min} \parent{\ell_2,\ell_3}}.
\end{align}

\begin{align}
\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}}
\end{align}
using Theorem 3.3 of \cite{bjork2013monte}.

On the other hand, theorem 3.2 and equation (3.54) states that
\begin{align}
\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{2 \kappa}  } ^{\frac{1}{2 \kappa}} \leq C_{5}^{CV} \parent{ \parent{\Delta t}^{\frac{1}{2}} + \Delta \tau }.
\end{align}

The weak rate for the discretisation error can be bounded by eq. (4.6):
\begin{align}
\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{}  }
= 
 \bigo{\parent{\Delta t}^2 + \parent{\Delta \tau}^2}
 = \bigo{2^{-2 \ell_1} + 2^{-2 \mathrm {min} \parent{\ell_2,\ell_3} }}
\end{align}

Due to the fact that the discretisation errors relating to $\tau$
depend on the minimum of $\ell_2$ and $\ell_3$, it seems natural
to assume $\ell_3=\ell_2$. This gives us a two-dimensional
Multi-Index Monte Carlo formulation with the following effective parameters
\begin{align}
\bar \beta &= [2,2]^T ,\\
\bar \gamma &= [1,1]^T , \\
\bar s &= [2^{-1},1], \\
\bar w & = [2,2].
\end{align}

On the other hand, one may set $\ell_1=\ell_2=\ell_3=\ell$
and arrive at a one-dimensional MLMC formulation with a weak rate of $2$, strong rate of $1$, and computational cost per realisation of $W_\ell=2^{2\ell}$. In the sequel, we will present the results of these two variations for comparison.

\section{Multi-Index Monte Carlo}

\section{Multi-Level Monte Carlo}

\bibliographystyle{agsm}
\bibliography{references}

\appendix

\end{document}