### 2016-05-06 tests/HJM/manuscript.tex
plots made for manu

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{mathrsfs }
\usepackage[section]{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{patchplots}
%\usepackage{subfigure}
\usepackage{subcaption}

\usepackage{hyperref}
\usepackage{comment}
\usepackage{multirow}

%line suggested by my compiler (Fabian)
\pgfplotsset{compat=1.9}

% put figures at the end of the document, while I am writing is very annoying to have the figures in the middle of the text! 
% \usepackage[nomarkers,figuresonly]{endfloat}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\input{macros.tex}

\title[MIMC and MLMC HJM]{Multi-Index and Multi-Level Monte Carlo Evaluation of HJM Models}
\author{Juho H\"app\"ol\"a}
\address{\kaustaddress}
\email{juho.happola@iki.fi}

\date{\today}    
                                    % Activate to display a given date or no date
\begin{document}

\begin{abstract}
Notes of MIMC and MLMC evaluation of HJM models.
\end{abstract}

\maketitle
%\tableofcontents

\section{HJM model in Fourier space}

Let us focus on the following HJM type SDE:
\begin{align}
d f \parent{t,\tau} &= \alpha \parent{t,\tau} dt + \beta \parent{t,\tau} d \overline W \parent t \\
f \parent{0,\tau} &= f_0 \parent \tau ,
\end{align}
with the diffusion term given as the infinite-dimensional extension of
eq. (1.1) in \cite{bjork2013monte}:
\begin{align}
%\beta \parent{t, \tau} d \overline W \parent t = \ssum{k=0}{\infty} c_n \parent{\sin %\parent{\ffreq{k} \tau} + \cos \parent{\ffreq{k} \tau} dW_n \parent t , 
\beta \parent{t, \tau} d \overline W \parent t = \ssum{k=0}{\infty}
c_n \parent{ \sin \parent{\ffreq k \tau} + \cos \parent{\ffreq k \tau} }
\end{align}
with $W_n$ independent Brownian motions and
\begin{align*}
\ffreq k \equiv \frac{k \pi }{L}
\end{align*}
with $L>>1$. This gives the following Fourier decomposition for the covariance
of increments:
\begin{align}
&\expp{ \beta \parent{t, \tau_1} d \overline W \parent t \beta \parent{t, \tau_2} d \overline W \parent t}
\\
= & \ssum{k=0}{\infty} c_k^2 \parent{ \cos \parent{\ffreq k \parent{\tau_1-\tau_2}}}
\end{align}

To keep the exponential HJM model risk neutral, we fix the drift of the equation as
\begin{align*}
\alpha \parent{t, \tau} 
=&
c_0^2 \parent{\tau-t}
\\
&+
\ssum{k=1}{\infty }
\frac{c_k^2}{k}
\sin \parent{\ffreq k \tau } \parent{\cos \parent{\ffreq q t} -  \sin \parent{\ffreq k t}}
\\
&+
\ssum{k=1}{\infty }
\frac{c_k^2}{k}
\cos \parent{ \ffreq k \tau } \parent{\cos \parent{\ffreq k t} -  \sin \parent{\ffreq k t}}
\\
&+
\ssum{n=1}{\infty }
\frac{c_n^2}{n}
\cos \parent{2 \ffreq k \tau }.
\end{align*}

The solution to the SDE can be written as
\begin{align*}
f \parent{t,\tau} - f_0 \parent \tau
= \tilde f \parent{t,\tau}
+ \ssum{n=1}{N} b_n \parent t \cos \parent{\frac{n \pi \tau}{L}} + a_n \parent t \sin \parent{\frac{n \pi \tau}{L}},
\end{align*}
with
\begin{align}
\tilde f \parent{t,\tau} =& c_0^2 t \parent{\tau-\frac{t}{2}},
\\
a_k \parent{t} = b_k \parent t =& \frac{c_k^2}{\ffreq k^2}\parent{\cos \parent{\ffreq k t} + \sin \parent{\ffreq k t}-1 } - \frac{\mathbf{1}_{\frac{k}{2} \in \mathbb Z_+}  c_j^2  t}{\ffreq k}  + c_k W_k \parent{t} .
\end{align}

The solution above lends itself to approximate solutions to the quantity of interest.
First, the discount factor can be approximated as:
\begin{align*}
&\int_0^{T} f \parent{s,s} ds
\\
\approx& \Delta t \ssum{n=1}{N_t} \parent{ f_0 \parent{t_n} + \tilde f \parent{t_n, t_n} +
\ssum{k=1}{N_f}
b_k \parent {t_n} \cos \parent{ \ffreq k t_n} + a_k \parent {t_n} \sin \parent{\ffreq k t_n}
}
\\
=& F_{N_t,N_f}
,
\end{align*}
with $t_n = \frac{nT}{N_t}$.
This approximation gives the following approximation error
in $N_f$.

The natural next step is, of coures, estimating the error in approximating the
integral by $F_{N_t,N_f}$. Firstly, there is the case of the frequency cutoff:
\begin{align*}
&\expp{ \parent{\int_0^T f \parent{s,s} ds - \lim_{N_t \rightarrow \infty} F_{N_t,N_f}}^2}
\\
=& \expp{\int_0^T \ssum{k=N_f+1}{\infty} \parent{b_k \parent{t_n} \cos \parent{\ffreq k t_n} + a_k \parent{t_n} \sin \parent{\ffreq k t_n}} ds}
\\
\approx &
\expp{\parent{\int_0^T \ssum{k=N_f+1}{\infty} c_k W_k \parent{t} \parent{\sin \parent{\ffreq k t} + \cos \parent{\ffreq k t} }}^2}
\\
= &
\expp{\parent{ \ssum{k=N_f+1}{\infty} \int_0^T c_k W_k \parent{t} \parent{\sin \parent{\ffreq k t} + \cos \parent{\ffreq k t} } ds}^2}
\\
= & \bigo{\ssum{k=N_f+1}{\infty} \parent{\frac{c_k}{k}}^2}.
\end{align*}
The bound for the weak error can be obtained using Jensen's inequality. Setting as a test
case the temporal correlation structure as
$
\rho \parent{\tau_1, \tau_2} \propto \expf{ - \kappa \absval{\tau_1-\tau2}}
$
it follows that $c^2_k =\frac{\kappa}{L} \frac{\parent{-1}^k\parent{1-\expf{-\absval{\kappa L}}}}{\kappa^2 + \parent{\ffreq k}^2}$. Using this as an example, we plot the empirical strong and weak convergence rates in images \ref{img:cutoff1weak} and \ref{img:cutoff1strong} respectively.

As for the time discretisation error, we are concerned with a rectangle quadrature
error, giving us weak and strong error rate of $1$ and $2$ respectively.

Similarly, one may estimate the underlying part of the 
payoff functional
\begin{align*}
&\int_{\tau_1}^{\tau_2}
f \parent{T,\tau} d\tau 
\\
\approx & \int_{\tau_1}^{\tau_2}
f_0 \parent \tau + \tilde f \parent{t_n, t_n} +
\ssum{k=1}{N_f}
b_k \parent T \cos \parent{\ffreq k \tau} + a_k \parent {T} \sin \parent{\ffreq k \tau}
d \tau.
\\
= & \int_{\tau_1}^{\tau_2}
f_0 \parent \tau  + \frac{c_0 T}{2} \parent{\tau_2^2 - \tau_2 +\tau_1 - \tau_1^2} d \tau
\\
&+
\ssum{k=1}{N_f} 
\frac{b_k \parent T}{\ffreq k } 
\parent{\sin \parent{\ffreq k \tau_2}  - \sin \parent{\ffreq k \tau_1} }
\\
&-
\ssum{k=1}{N_f} 
\frac{a_k \parent T}{\ffreq k} 
\parent{\cos \parent{\ffreq k \tau_2}  - \cos \parent{\ffreq k \tau_1} }
\\
=& \Psi_{N_f} .
\end{align*}
The corresponding empirical rates for the approximations of the quantity of interest
\begin{align*}
\mathcal G \parent f =
\expp{\expf{-\int_0^{t_{T}} f \parent{s,s} ds}\expf{-\int_{\tau_1}^{\tau_2} f \parent{T,\tau} d\tau}}
\approx
\expp{
F_{N_t,N_f}
\Psi_{N_f} }
\end{align*}
are presented in fig. \ref{fig:empInfDimRates}

Setting $\ell = \parent{\ell_1 , \ell_2}$, we may define a Monte
Carlo estimator using $m$ independent realisations of $\Psi_{N_f}$ and $F_{N_t,N_f}$:
\begin{align}
\mathcal A_{\ell_1, \ell_2} 
\equiv
\ssum{m=1}{M}
\frac{F_{C_1 2^{\ell_1}+1,C_2 2^{\ell_2}} \Psi_{C_2 2^{\ell_2} }\parent m}{M}.
\end{align}

Similarly, we may extend the above to a MLMC estimator as
\begin{align*}
\mathcal{A}_{ML} =& \ssum{m=0}{M_0} \frac{F_{2 C_1 ,C_2 } \Psi_{C_2 }\parent m}{M_0}
\\
&+
\ssum{\ell_1=1}{L}
\ssum{m=0}{M_{\ell_1}} \frac{\parent{ F_{C_1 2^{\ell_1}+1,C_2 2^{\ell_1}} \Psi_{C_2 2^{\ell_1} } - F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_1-1}} \Psi_{C_2 2^{\ell_1} }} \parent m }{M_{\ell_1}},
\\
\equiv& \ssum{\ell_1=0}{L} \ssum{m=0}{M_{\ell_1}} \frac{\Delta_{\ell_1} \parent m}{M_{\ell_1}}
\end{align*}
and, into a MIMC estimator through defining the appropriate two-dimensional difference operators $\Delta_{\ell_1,\ell_2}$ and a downward-closed index-set $L_K=\sset{\parent{\ell_1,\ell_2}\in \mathbb Z_+^2: I \parent{\ell_1,\ell_2}<K}$:
\begin{align*}
\mathcal A_{MI} =& \ssum{\ell \in L_K}{} \ssum{m=1}{M_{\ell}} \frac{\Delta_{\ell_1,\ell_2} \parent m}{M_\ell}
\\
\Delta_{\ell_1,\ell_2}
\equiv&
F_{C_1 2^{\ell_1}+1,C_2 2^{\ell_2}} \Psi_{C_2 2^{\ell_2} } - \mathbf{1}_{\ell_1>1} F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_2-1}} \Psi_{C_2 2^{\ell_2-1} }
\\
& -\mathbf{1}_{\ell_2>1} F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_2}} \Psi_{C_2 2^{\ell_2} }
+ \mathbf{1}_{\ell_1,\ell_2>1} F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_2-1}} \Psi_{C_2 2^{\ell_2-1} }.
\end{align*}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{nfstrong.pdf}
        %\caption{A gull}
        \label{fig:gull}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{nfweak.pdf}
        %\caption{A tiger}
        \label{fig:tiger}
    \end{subfigure}
    \\
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{ntstrong.pdf}
        %\caption{A mouse}
        \label{fig:mouse}
    \end{subfigure}
        \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{ntweak.pdf}
        %\caption{A mouse}
        \label{fig:mouse}
    \end{subfigure}
    \caption{\label{fig:empInfDimRates}Empirical strong and weak convergence rate for
    the $N_t$ and $N_f$.}
\end{figure}

\bibliographystyle{agsm}
\bibliography{references}

\appendix

\end{document}